---
title: "Data visualisation"
author: "Marina Espinasse"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    toc: true
    toc_depth: 2
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: pygments
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(here)
library(sf)
library(ggthemes)
library(RColorBrewer)
```

## Loading data and shapefiles

```{r}
covid <- read.csv(here("output", "covid19_monthly_mean_n_total.csv"))

covid$mcp_code <- as.character(covid$mcp_code)
```

```{r, message=F}
map <- st_read(here("data", "belgium_shape", "Apn_AdMu.shp"))
# Projected CRS: ETRS89 / Belgian Lambert 2008
```



# Preparation of the data

In the MAP dataset, we are interested in the municipality code,AdMuKey and the geometry. The other columns of interest (province and covid19 cases) I will add from the covid19 table (by left joining it to map dataframe).

```{r}
map_prep <- map |>
  select(c(AdMuKey, Stat_Area:geometry)) |>
  rename(mcp_code = AdMuKey) |>
  janitor::clean_names() |>
  left_join(covid, by = "mcp_code") |>
  mutate(month_char = recode(
    month, "January", "February",
    "March", "April",
    "May", "June", "July",
    "August", "September", "October", "November",
    "December"
  ))
```

# Mapping total monthly COVID-19 cases

```{r}
breaks1 <- seq(0, 14000, by = 1000)
```


```{r}
map1 <- ggplot(map_prep) +
  geom_sf(aes(fill = total_monthly), lwd = 0.1) +
  scale_fill_gradientn(
    colours = brewer.pal(9, "PuRd"),
    breaks = c(0, 10, 100, 1000, 10000),
    labels = c(0, 10, 100, 1000, 10000),
    trans = "log1p"
  ) +
  labs(fill = "Total COVID-19 \ncases per month") +
  facet_wrap(. ~ fct_reorder(month_char, month, min)) +
  theme_void() +
  theme(strip.text = element_text(face = "bold", size = 11))
```


```{r}
ggsave(
  filename = "./figs/monthly_total_covid_mcp.pdf", plot = map1,
  height = 20, width = 30, units = "cm"
)
```

# Cases of COVID-19 per municipality and month

```{r}
cases_province <- covid |>
  group_by(province, month) |>
  summarise(
    total = sum(total_monthly),
    mean = mean(mean_monthly)
  ) |>
  mutate(month_char = recode(
    month, "January", "February",
    "March", "April",
    "May", "June", "July",
    "August", "September", "October", "November",
    "December"
  )) |>
  mutate(province = replace(province, province == "OostVlaanderen", "Oost-Vlaanderen")) |>
  mutate(province = replace(province, province == "WestVlaanderen", "West-Vlaanderen")) |>
  mutate(province = replace(province, province == "VlaamsBrabant", "Vlaams-Brabant")) |>
  mutate(province = replace(province, province == "BrabantWallon", "Brabant-Wallon"))
```




```{r}
plot2 <- ggplot(
  data = cases_province,
  aes(
    x = fct_reorder(month_char, month, min),
    y = total, group = 1
  )
) +
  geom_line(alpha = 0.8, color = "deeppink1", linewidth = 1.5) +
  geom_point(color = "deeppink3", aes(size = total)) +
  facet_wrap(province ~ .) +
  theme_bw() +
  theme(
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text = element_text(size = 12, face = "bold"),
    axis.text.x = element_text(angle = 90),
    legend.title = element_blank()
  ) +
  labs(y = "Total cases", x = "", fill = "Total")
```

```{r}
ggsave("./figs/monthly_total_covid_province.pdf",
  plot = plot2,
  height = 20, width = 25, units = "cm"
)
```


# Incidence of COVID-19 in 2021


```{r}
pop <- read.delim(here("data", "TF_SOC_POP_STRUCT_2020.txt"), sep = ";")
```

Clean the population dataset and calculate incidence per 10000 citizens

```{r}
pop_prep <- pop |> 
  select(c(CD_REFNIS,CD_SEX:MS_POPULATION )) |> 
  group_by(CD_REFNIS) |> 
  summarize(pop_tot = sum(MS_POPULATION)) |> 
  rename(mcp_code = CD_REFNIS) |> 
  mutate(mcp_code = as.character(mcp_code))
```

Join map data and population data:

```{r}
map_incidence <- map_prep |> 
  left_join(pop_prep, by = "mcp_code") |> 
  mutate(incidence = total_monthly/ pop_tot * 10^4) #incidence per 10000 citizens
```


Save the table of COVID-19 incidences per month

```{r}
incidence_dataset <- st_set_geometry(map_incidence, NULL) |> 
  select(c(mcp_code, mcp_name:incidence))
```

Save this table

```{r}
write.csv(incidence_dataset, file = "./output/covid19_incidence_month.csv", row.names = F)
```



Create a map

```{r}
map2 <- ggplot(map_incidence) +
  geom_sf(aes(fill = incidence), size = 0.2) +
  scale_fill_gradientn(
    colours = brewer.pal(9, "PuRd"),
    breaks = c(0, 10, 100,500, 1000),
    labels = c(0, 10, 100, 500, 1000),
    trans = "log1p"
  ) +
  labs(fill = "COVID-19 incidence") +
  facet_wrap(. ~ fct_reorder(month_char, month, min)) +
  theme_void() +
  theme(strip.text = element_text(face = "bold", size = 11),
        legend.title = element_text(face = "bold", size = 10))

```


```{r}
ggsave("./figs/incidence_covid_mcp.pdf",
  plot = map2,
  height = 25, width = 36, units = "cm"
)
```