---
title: "Preparation of crossbasis variables of pollutaion data for INLA models"
author: "Marina Espinasse"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    toc: true
    toc_depth: 2
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: pygments
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r}
library(tidyverse)
library(here)
library(INLA)
library(tsModel)
library(sf)
library(spdep)
library(sp)
library(rgdal)
library(dlnm)
```



In this rmd file, I will prepare lags of the covariates (pollution and vaccination), and an adjacency matrix for inla models. Complete dataset for inla modelling will be saved in the "output" folder

# Data

Load master file on COVID-19 cases, population, and pollution per week.


```{r}
data <- read.csv(here("output", "master_dataset.csv"))
```

Arrange data by year and weeknumber within each municipality

```{r}
data_sort <- data |> 
  arrange(year, week_num)
```


# Create lagged variables

Set maximum lag to 6 weeks

```{r}
nlag = 6
```

For creating lags of pollution data and vaccination data, I will use `tsModel` package.

We observed strong positive correlation between pm10 and pm2.5 (0.9), and between BC and NO2 (0.8).
It is possible that one of pollutants in each of these pairs will not be included in the final analysis, but we first can test them individually by comparing DIC.

```{r}
lag_ozone <- tsModel::Lag(data_sort$pollution_ozone_wk, group = data_sort$mcp_code, k = 0:nlag)
lag_no2 <- tsModel::Lag(data_sort$pollution_no2_wk, group = data_sort$mcp_code, k = 0:nlag)
lag_pm25 <- tsModel::Lag(data_sort$pollution_pm25_wk, group = data_sort$mcp_code, k = 0:nlag)
lag_pm10 <- tsModel::Lag(data_sort$pollution_pm10_wk, group = data_sort$mcp_code, k = 0:nlag)
lag_bc <- tsModel::Lag(data_sort$pollution_bc_wk, group = data_sort$mcp_code, k = 0:nlag)
lag_vaccine <-tsModel::Lag(data_sort$cumul_week, group = data_sort$mcp_code, k = 0:nlag)
```

We have to remove the  weeks  of year 2020, because it will only have NA for covid19 response data.
We needed year 2020 (the weeks 47 to 52) only to create 6-weeks lags in the pollution variables.

For vaccine data, I do the same, but since we do not have data from 2020 for vaccine, it will have 3486 (581*6) rows with missing data. But there are additional 17 missing data for 17 municipalities that did not have vaccination data in 2021 (see 01_data_preparation.Rmd file for the list of these vaccines).

```{r}
colSums(is.na(lag_ozone))
#lag_ozone <- lag_ozone[!(data_sort$week_num < 47 & data_sort$year == 2020),]
lag_ozone <- lag_ozone[data_sort$year > 2020,]
lag_no2 <- lag_no2[data_sort$year > 2020, ]
lag_bc <- lag_bc[data_sort$year > 2020, ]

lag_pm25 <- lag_pm25[data_sort$year > 2020, ]
lag_pm10 <- lag_pm10[data_sort$year > 2020, ]


lag_vaccine <- lag_vaccine[data_sort$year > 2020, ]

# remove weeks year 2020 from the main dataset
data_prep <- data_sort[data_sort$year > 2020,]
```


# Defining cross-basis functions

For know, we select simple functions. 1) for response-exposure dimension, we use natural splines with 2 knots (and 2+1=3 df), knots are equally spaced; 2) for lag-response dimension, we use 1 knot with (1 + 1 + intercept = 3 df). Total number of variables = 3*3=9.

```{r}
basis_ozone <- crossbasis(lag_ozone,
                    argvar = list(fun = "ns", knots = equalknots(data_prep$pollution_ozone_wk, 2)),
                    arglag = list(fun = "ns", knots = nlag/2))

basis_no2 <-crossbasis(lag_no2,
                    argvar = list(fun = "ns", knots = equalknots(data_prep$pollution_no2, 2)),
                    arglag = list(fun = "ns", knots = nlag/2))
  
  
basis_pm25 <- crossbasis(lag_pm25,
                    argvar = list(fun = "ns", knots = equalknots(data_prep$pollution_pm25, 2)),
                    arglag = list(fun = "ns", knots = nlag/2))
  
basis_pm10 <- crossbasis(lag_pm10,
                    argvar = list(fun = "ns", knots = equalknots(data_prep$pollution_pm10, 2)),
                    arglag = list(fun = "ns", knots = nlag/2))
  
basis_bc <-crossbasis(lag_bc,
                    argvar = list(fun = "ns", knots = equalknots(data_prep$pollution_bc, 2)),
                    arglag = list(fun = "ns", knots = nlag/2))

basis_vaccine <- crossbasis(lag_vaccine,
                    argvar = list(fun = "ns", knots = equalknots(data_prep$cumul_week, 2)),
                    arglag = list(fun = "ns", knots = nlag/2))
```


Column names for crossbasis function are the same for crossbasis of ozone, pm25, bc and other variables. Here I create unique column names.

```{r}
colnames(basis_ozone) = paste0("basis_ozone.", colnames(basis_ozone))
colnames(basis_no2) = paste0("basis_no2.", colnames(basis_no2))
colnames(basis_pm25) = paste0("basis_pm25.", colnames(basis_pm25))
colnames(basis_pm10) = paste0("basis_pm10.", colnames(basis_pm10))
colnames(basis_bc) = paste0("basis_bc.", colnames(basis_bc))
colnames(basis_vaccine) = paste0("basis_vaccine.", colnames(basis_vaccine))
```





Define penalized complexity prior for spatial random effect and for temporal effect

```{r}
precision.prior <- list(prec = list(prior = "pc.prec", param = c(0.5, 0.01)))
```

INLA model function. Note that in the control.predictor we set link=1, this means we want a logit link function.





# Create adjacency matrix and function for INLA models


Municipalities in the map are sorted in the same way as in the master data, but to be sure, I will connect the map file and master data by creating a municipality id (chunk 16).

```{r, message=F}
# load shape file for Belgium
map <- readOGR("./data/belgium_shape/Apn_AdMu.shp")

neig.map <- poly2nb(map,row.names = map$AdMuKey)
```

Plot adjacency matrix and save the neighbors plots
```{r, echo=FALSE}
tiff(file = "./figs/adjacency_mat.tiff",width = 7, height = 6, units = 'in', res = 250)
plot(map, border = grey(0.5))   #Don't close the graph
Coords <- coordinates(map)
# Connect neighbours with a line
plot(neig.map, 
     coords = Coords, 
     add = TRUE,
     pch = 20,
     lwd = 1,
     col = 2)
dev.off()
```

Convert the spatial object of neighbors to an inla object again

```{r}
adj.file <- "output/adjacency.mat"
if (!file.exists(adj.file)) nb2INLA(adj.file, neig.map)

```

```{r}
mymodel <- function(formula, data = df, family = "nbinomial", config = FALSE)

  {
  model <- inla(formula = formula, data = data, family = family, offset = log(E),
       control.inla = list(strategy = 'adaptive'), 
       control.compute = list(dic = TRUE, config = config, 
                              cpo = TRUE, return.marginals = FALSE),
       control.fixed = list(correlation.matrix = TRUE, 
                            prec.intercept = 1, prec = 1),
       control.predictor = list(link = 1, compute = TRUE), 
       verbose = FALSE)
  model <- inla.rerun(model)
  return(model)
}
```

# Additional variables for a final dataframe

Additional variables are total number of weeks (52), total number of municipalities (581), and provinces (11)

```{r}
# total number of weeks
nweeks <- length(unique(data_prep$week))
#total number of muncipalities
nmcp <- length(unique(data_prep$mcp_code))
# total number of provinces
nprov <- length(unique(data_prep$province))
```

Here I create a municipality identifier. I take the municipalities' codes as they are in the map file, turn them into a sequence from 1 to 581, and add this sequence to my dataset. This way, I will be sure that municipality ID in the adjacency matrix and in my dataset are identical.

```{r}
#index for municipality (run chunk 15 first!)
mcp_index <- data.frame(
  mcp_code = map$AdMuKey,
  mcp_id = seq(1, length(map$AdMuKey))
)
mcp_index$mcp_code <- as.numeric(mcp_index$mcp_code)

data_prep <- left_join(data_prep, mcp_index, by = "mcp_code")
```


Create final variables for INLA models 

```{r}
Y <- data_prep$cases_per_week  #observed cases of COVID-19 per week
N <- length(Y)  #total length of the dataset
E <- data_prep$pop_tot/10^3  #for the incidence per 1000 population
T1 <- data_prep$week_num  #week indicator = week number
S1 <- data_prep$mcp_id  #municipality indicator (1 to 581)
```

Create a dataframe for model testing

```{r}
df <- data.frame(Y, E, T1,S1)
```


